The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
More tune-ups to section 4.6 and 4.7
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@714 4525493e-7705-40b1-a816-d608a930855b

\subsection{DAS Mapping DB}
The Mapping DB represents DAS {\it registration} service. It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

\subsection{DAS queries}
All DAS queries were expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We used a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on the analysis of daily work operations of our users, see \cite{DBS07}.
Each key represents a DAS record. Every record was stored into the cache
in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
we were able to store arbitrary structure of DAS records, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
set of attributes representing JSON structure. For instance a document
\begin{verbatim}
{'block': {'name':'abc', 'size': 1, 'replica': [...]}}
\end{verbatim}
associated with a \verb+block+ key and therefore had the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes were based 
on a set of meta-data information provided by data-services. Our users 
where able to specify a context in their queries by using appropriate 
attribute(s). For example, for query \verb+block+ we didn't perform a string 
matching to the ``block'' string, but rather look-up a block records, while query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matched names. The numerical values, conditions, groupping were
easy to express too. For example, \verb+run > 10 and run < 100+ query 
was used to get all run records for provided range of run numbers. 
At the same time, to complement the strength of the queries, we do plan 
to index dynamically DAS records and add a keyword-based 
search later.

All input queries were transformed into MongoDB query syntax, 
who was used internally in DAS. It represents a set of selection 
keys and conditions as a JSON document. For example, a query
\verb+dataset file=abc run>10+
was transformed into 
\begin{verbatim}
{"fields": {"dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries were stored into Analytics DB similar to other DAS 
records and used for further analysis.
%Such syntax was perfect fit for us, since it allows to store all queries
%as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
%MongoDB query syntax is very reach. It supports advanced queries with 
%boolean expressions, conditional operators, regular expressions, etc. providing
%a real power for our end-users.

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB became an authorative
source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provided information about CMS block meta-data. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas. Upon a user request to 
find a block meta-data, we were able to identify appropriate set of 
DBS and PhEDEx APIs, retrieve data from them, re-format their output 
into DAS notations and store results into DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ refers to api call records stored in Analytics DB,
\verb+block+ represents an aggregation block meta-data information from DBS
and PhEDEx systems and \verb+das+ contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.

>>>>

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB represents DAS {\it registration} service. It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

\subsection{DAS queries}
All DAS queries were expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We used a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on the analysis of daily work operations of our users, see \cite{DBS07}.
Each key represents a DAS record. Every record was stored into the cache
in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
we were able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key had a 
set of attributes representing JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
associated with a \verb+block+ key who had the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record and therefore
a nested structures were allowed, e.g. \verb+block.replica.size+. Our users 
where able to specify a context of their queries by using appropriate 
attribute(s). For example, query \verb+block+ represents search for all
block records rather then ``block'' string, while query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matched names. The numerical values, conditions, groupping were
easy to express too, e.g. \verb+run > 10 and run < 100+ query 
was used to get all run records for provided range of run numbers. 
To complement the strength of the queries, we do plan to index 
dynamically DAS records and add a keyword-based 
search within record content later.

All input queries were transformed into MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run>10+
was represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries were stored into Analytics DB similar to other DAS 
records and used for further analysis.
%Such syntax was perfect fit for us, since it allows to store all queries
%as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
%MongoDB query syntax is very reach. It supports advanced queries with 
%boolean expressions, conditional operators, regular expressions, etc. providing
%a real power for our end-users.

\subsection{Data-services and aggregation}
As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
were provided information about CMS blocks. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas. Upon a user request to 
find a block meta-data, we were able to identify appropriate set of 
DBS and PhEDEx APIs, retrieve data, re-format their output 
into DAS notations and store results into DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ part refers to api call records stored in Analytics DB,
the \verb+block+ part represents an aggregation block meta-data information from DBS
and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.
